---
title: "dpi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dpi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dpi)
```

# Background

This package aims to provide a minimalist and uniform interface for retrieving 
data products and the associated metadata no matter where the data product is 
housed. Currently, s3 and labkey are two types of remote repositories for 
housing data products. 

This packages leverages the `pins` package for accessing data. As such, it 
leverage the concept of `boards` for organizing data into **data spaces**. As 
an example, a team working on multiple clinical studies, may consider defining 
a data space or `board` for each study, each of which defined by a governance 
policy. In this example, a `board` is not dissimilar to the 
idea of a directory or folder. A board can be instantiated on different data 
platforms (e.g. AWS s3) supported by `pins` or its extensions. 

As different data platforms (e.g. s3 vs. labkey) have their own specific url and
ways of authentication, different parameters need to be set up for how 
connection to the remote data repository is to be established. Once the 
connection is established, however, all other interactions (e.g. list, get, 
version, etc.) are done using the same set of functions.  With that in mind, 
the steps outlined below are divided into 

1. Setting up the connection
2. Interacting with the data product

# Set up connection

Setting up connection to data products involves setting

1. Board parameters: parameters relating to the remote location of the data 
product
2. Credential parameters: parameters checked by the remote location of the data 
when trying to access

**NOTE:** prior to proceeding further verify that you have working credentials 
for the repository you need to access.


## Board parameters

Currently data products housed on either Labkey or AWS s3 are supported. 

- For s3, you need to know the bucket name and the region.
- For Labkey, you need to know the url of platform and specific folder.


## Format board parameters

For S3, you need to provide an alias that signifies the data space this data 
product could belong to. 

```{r, eval=FALSE}
board_params_s3 <- board_params_set_s3(board_alias = "an_alias_for_data_board",
                                 bucket_name = "bucket_name",
                                 region = "us-east-1")
```

The Labkey parameters for a data product will have pattern as given below

```{r, eval=FALSE}
board_params_labkey <- board_params_set_labkey(board_alias = "labkey",
                                               url = 'https:url_to_labkey/labkey',
                                               folder = '/project_folder/subfolder')
```



## Format your credentials

This can be done with `cred_set_aws` or `cred_set_labkey` functions. The 
cred_set function signatures informs you of the type of parameters needed.

**NOTE:** For security reasons, avoid putting your credentials inside your 
script. It is highly recommended to establish a habit of using environment 
variables in conjunction with secret management packages like `keyring`. Below,
exemplifies this pattern, where the env variables are set using 
`Sys.setenv("AWS_KEY" = "xxxx")` in console, or using `keyring`,
`Sys.setenv("AWS_KEY" = keyring::key_get(service = "AWS_KEY")`. Similarly, 
"AWS_SCRET" can be set as an environment variable.


The signature for creds for AWS s3 is takes two parameters.

```{r, eval=FALSE}
aws_creds <- creds_set_aws(key = Sys.getenv("AWS_KEY"),
                           secret = Sys.getenv("AWS_SECRET"))
```


If the data board was to be set up on Labkey, the signature for creds only 
takes one parameter `api_key`
```{r, eval=FALSE}
labkey_creds <- creds_set_labkey(api_key = Sys.getenv("LABKEY_API_KEY"))

```


# Connect and interact with data product

With the connection parameters established, interaction with the data product 
use the same set of function and format no matter where the data is housed.

## Connect to the board

Then we can connect to the board.

```{r, eval=FALSE}
dp_connect(board_params = board_params_s3, creds = aws_creds)
```

## Work with the board

List what data products available on this board we just connected to. Note we 
need to connect first in order to list.

```{r, eval=FALSE}
dp_list(board_params = board_params_s3)
```

## Retrieve data

Load data into the working environment, e.g. as object `dp`.

```{r, eval=FALSE}
dp <- dp_get(data_name = "dp-studyid_branchid", 
            board_params = board_params_s3, 
            version = "c5a51c3")
```
